### 2013-02-15 doc/icml2013/algo.tex
many changes

In this section, we describe our Spectral Experts algorithm
for estimating model parameters $\theta = (\pi, B, \sigma^2)$.
Our algorithm consists of two steps:
(i) low-rank regression to estimate certain symmetric tensors;
and (ii) tensor factorzation to recover the parameters.
% Motivate how we can recover $B$ from regression by lower order moments
%ideally from lower order moments.
%Conventionally, when using the method of moments, the
%moments of observed variables are expressed algebraically in terms of
%the model parameters. Using these equations, we attempt to solve for the
%model parameters using empirical estimates of the moments.

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
independent of $x$.
We can average over this randomness by defining
the average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
plus some noise $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}.
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$.

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
%y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \underbrace{\innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}}}_{\eqdef \eta_2(x)} + (\epsilon^2 - \sigma^2) \label{eq:y2}, \\
\end{align}
Again, we have expressed $y^2$ has a linear function of $[x\tp{2}, 1]$
with regression coefficients $[M_2, \sigma^2]$, plus noise.
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

Performing regression yields a consistent estimate of $[M_2, \sigma^2]$,
but does not identify all the parameters $\theta$.
In particular, $B$ is only identified up to rotation:
if $B = [\beta_1 \mid \cdots \mid \beta_k]$ satisfies
$B \diag(\pi) B^\top = M_2$, then $(B Q) \diag(\pi) (Q B^\top)$
for any orthogonal matrix $Q$.
%Regression with the second order moments would give us $M_2 = \E[\beta_h
%\tp{2}] = \sum_h \pi_h \beta_h\tp{2}$ which can identify $\pi$ and the
%subspace spanned the $\beta_h$, but not the $\beta_h$ themselves. To see
%this, note that the eigen-decomposition of $M_2$ is not unique.
%\todo{explain better}.

%%% third moment
Let us now proceed to the third moment with the hope that it will
provide additional information.
We can write $y^3$ as a linear function of $x\tp{3}$ plus zero mean noise:
\begin{align}
y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
\eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
+ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
&+ 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
+ \epsilon^3, \nonumber
\end{align}
where $\E[\eta_3(x) \mid x] = 0$.
Performing regression on our data yields estimates $M_3$ and $\sigma^2 M_1$.
It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
all the parameters.

%We note that the eigendecomposition of $M_2$ is not sufficient to
%recover the $\beta_k$ because they are not necessarily orthogonal to
%each other.

% Full algorithm
Now we are ready to state our full algorithm, which we call Spectral Experts.
The full algorithm is described in \algorithmref{sec:algo}.
First, we perform two regressions according to \refeqn{y2}
and \refeqn{y3} to recover the compound parameters $M_2$ and $M_3$.
The first regression \refeqn{estimateM2}
produces estimates of $\hat M_2$ and $\hat\sigma^2$;
the second regression \refeqn{estimateM3} produces $\hat M_3$ (we ignore $\hat S_1$).
Since $M_2$ and $M_3$ have low rank structure,
we use nuclear norm regularization \cite{Tomioka2011,NegahbanWainwright2009}
to exploit this structure.
The regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
are set to $O(\frac{1}{\sqrt{n}})$.
See \sectionref{sec:regression} for more details.

Second, given a matrix $M_2$ and tensor $M_3$ of the form in \refeqn{Mp},
\citet{AnandkumarGeHsu2012} showed that due to symmetry of $M_2$ and $M_3$,
we can efficiently perform tensor factorization of the whitened
version of $M_3$.  The eigenvalues and eigenvectors
directly yield the parameters $\pi$ and
$\{\beta_h\}$.
See \sectionref{sec:tensor-power} for details.
%\citet{AnandkumarGeHsu2012} showed that given $M_2$ and $M_3$
%of the form in \refeqn{Mp},
%the robust tensor power method can be employed to recover $\pi$ and
%$\{\beta_h\}$.

@@

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Data $\mathcal{D} = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k], \hat\sigma^2)$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &[\hat M_2, \hat \sigma^2] = \arg\min_{M_2,\sigma^2} \label{eqn:estimateM2} \\
      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
       &[\hat M_3, \hat S_1] = \arg\min_{M_3,S_1} \label{eqn:estimateM3} \\
       &\quad\frac{1}{2n} \sum_{(x,y) \in \sD} (\innerp{M_3}{x\tp{3}} + \innerp{S_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
    \end{align}
    \STATE Estimate the coefficients $\{\beta_h\}$ using \textbf{tensor factorzation}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $W \in \Re^{d \times k}$ (such that $W^\top
      \hat M_2 W = I$) using SVD.
      \item [(b)] Compute the eigenvalues $\{\lambda_h\}_{h=1}^k$
      and eigenvectors $\{v_h\}_{h=1}^k$
      of the whitened tensor $M_3(W, W, W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\lambda_h^2}$
    and $\hat\beta_h = (W^{\top})^\dagger (\lambda_h v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

Our main theoretical result shows that Spectral Experts returns
consistent parameter estimates (see Appendix
\begin{theorem}[Convergence of Spectral Experts]
Let $\sD$ consist of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$.
Assume that $\E[[1,\unfold(x\tp{2})]\tp{2}] \succ 0$ 
and $\E[[x,\unfold(x\tp{3})]\tp{2}] \succ 0$.
Then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts}, with the columns appropriate permuted,
satisfies $\sigma^2 \le \epsilon$,
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\| \le \epsilon$,
for some error tolerance $\epsilon = O(\frac{kd^2}{n (\min_h \pi_h^*)})$.
\end{theorem}
\begin{proof}
Let $T^* = M_3^*(W^*, W^*, W^*)$.
By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
\end{proof}

% Method of moments interpretation
%Though we have motivated Spectral Experts using regression on powers
%of $y$ given powers of $x$,
%we can also think of the algorithm in terms of
%the method of moments.  Specifically,
%the optimality conditions for the regression of $M_2$ and $M_3$ are:
%\begin{align}
%\E[x\tp{2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)] = 0, \\
%\E[x\tp{3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2 \innerp{M_1}{x} - y^3)] = 0.
%\end{align}

@@

\subsection{Identifiability from moments}

The mixture of linear regressions model is identifiable subject to
permutations of the components\citationneeded. However, our approach
employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
these condition {\em quadratic} and {\em cubic} independence.

This condition trivially holds when the space $\mathcal{X}$ consists of
independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
places some constraints on non-linear featurizations. For example, the
common polynomial basis expansion, $1, x_1, x_1^2, \dots, x_1^p$ is
neither quadratically nor cubically independent because the product
space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$. In fact, the
number of unique terms in $x\tp{p}$ is only of the order of $p \times
d$, and hence we will not have enough equations to solve for $M_2$ or
$M_3$ unless we consider $p = d^2$ moments.

>>>>

In this section, we describe our Spectral Experts algorithm
for estimating model parameters $\theta = (\pi, B)$.
The algorithm consists of two steps:
(i) low-rank regression to estimate certain symmetric tensors;
and (ii) tensor factorzation to recover the parameters.
The two steps can be performed efficiently using
convex optimization and tensor power method, respectively.
% Motivate how we can recover $B$ from regression by lower order moments
%ideally from lower order moments.
%Conventionally, when using the method of moments, the
%moments of observed variables are expressed algebraically in terms of
%the model parameters. Using these equations, we attempt to solve for the
%model parameters using empirical estimates of the moments.

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$ where
$\epsilon \sim \normal{0}{\sigma^2}$, and $h$ is a random quantity
independent of $x$.
We can average over this randomness by defining
the average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with coefficients $M_1$
plus some noise $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{M_1 - \beta_h}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \normal{0}{\sigma^2}$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$
under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$.

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

Performing regression yields a consistent estimate of $M_2$,
but does not identify all the parameters $\theta$.
In particular, $B$ is only identified up to rotation:
if $B = [\beta_1 \mid \cdots \mid \beta_k]$ satisfies
$B \diag(\pi) B^\top = M_2$, then $(B Q) \diag(\pi) (Q B^\top)$
for any orthogonal matrix $Q$.

%%% third moment
Let us look to the third moment for additional information.
We can write $y^3$ as a linear function of $x\tp{3}$ with coefficients $M_3$
plus a bias $3 \sigma^2 \innerp{M_1}{x}$ that can be estimated from \refeqn{y1} and zero mean noise $\eta_3(x)$:
\begin{align}
y^3 &= \innerp{M_3}{x\tp{3}} + 3\sigma^2 \innerp{M_1}{x} + \eta_3(x), \label{eqn:y3} \\
\eta_3(x) &= \innerp{\beta_h\tp{3} - M_3}{x\tp{3}}
+ 3 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} \nonumber \\
&\quad + 3(\epsilon^2 \innerp{\beta_h}{x} - \sigma^2 \innerp{M_1}{x})
+ \epsilon^3. \nonumber
\end{align}
Performing regression yields estimates for $M_3$.
It turns out that knowledge of $M_2$ and $M_3$ are sufficient to recover
all the parameters.

% Full algorithm
Now we are ready to state our full algorithm, which we call Spectral Experts
(\algorithmref{sec:algo}).
First, we perform three regressions to recover the compound parameters
$M_1$ \refeqn{y1},
$M_2$ \refeqn{y2}, and
$M_3$ \refeqn{y3}.
Since $M_2$ and $M_3$ both only have rank $k$,
we can use nuclear norm regularization
\cite{Tomioka2011,NegahbanWainwright2009}
to exploit this low-rank structure and improve our compound parameter estimates.
In the algorithm, the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(2)}$
are set to $\frac{1}{\sqrt{n}}$.
The resulting semidefinite program is a standard one which has received
much attention in recent years.
We use a rather simple gradient-based approach,
where the nuclear norm is handled in closed form by taking an SVD
and soft-thresholding the eigenvalues \cite{donoho95soft,cai10soft}.

% Tensor factorization
Having estimated the compound parameters $M_2$ and $M_3$,
it remains to recover the original parameters $\theta$.
\citet{AnandkumarGeHsu2012} showed that for $M_2$ and $M_3$ of
the forms in \refeqn{Mp}, it is possible to efficiently accomplish this.
Specifically, we first compute a whitening matrix based on the SVD of $M_2$
and use that to construct a tensor $T = M_3(W, W, W)$ whose factors are orthogonal.
We can use the robust tensor power method to compute all the
eigenvalues and eigenvectors of $T$, from which it is easy to recover
the parameters $\pi$ and $\{\beta_h\}$.

% Unmixing
At a conceptual level, Spectral Experts bears some resemblence to the unmixing
algorithm for estimation of restricted PCFGs
\cite{hsu12identifiability}.
In that work, the observations (moments) provided a linear combination over
the compound parameters.  ``Unmixing'' involves solving for the compound
parameters by inverting a mixing matrix.  In this work,
each data point (appropriately transformed) provides a noisy slice of the
compound parameters; retrieving the compound parameters is more involved.

% Signal
The idea of performing low-rank regression on $y^2$ has been explored
in the context of signal recovery from magnitude measurements
\cite{candes11phaselift,ohlsson12phase}.
There, the actual observed response was $y^2$,
whereas in our case, we delibrately construct powers $y,y^2,y^3$
to identify the parameters.

@@

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Data $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$, $r = 1, 2, 3$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x\tp{2}} - y)^2, \nonumber \\
      &\hat M_2 = \arg\min_{M_2} \label{eqn:estimateM2} \\
      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2 + \lambda_n^{(2)} \|M_2\|_*, \nonumber \\
      &\hat M_3 = \arg\min_{M_3} \label{eqn:estimateM3} \\
      &\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2 + \lambda_n^{(3)} \|M_3\|_*. \nonumber
    \end{align}
    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorzation}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $W \in \Re^{d \times k}$ (such that $W^\top
      \hat M_2 W = I$) using SVD.
      \item [(b)] Compute the eigenvalues $\{a_h\}_{h=1}^k$
      and eigenvectors $\{v_h\}_{h=1}^k$
      of the whitened tensor $M_3(W, W, W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{a_h^2}$
    and $\hat\beta_h = (W^{\top})^\dagger (a_h v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theoretical results}

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\unfold(x\tp{2})\tp{2}] \succ 0$ 
and $\E[\unfold(x\tp{3})\tp{2}] \succ 0$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
There exists universal constants $c_1,c_2$ such that
if the number of samples is
$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
\end{theorem}
\begin{proof}
Let $T^* = M_3^*(W^*, W^*, W^*)$.
By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
Need to convert error from $\epsilon$
hit by $(W^\top)^\dagger$
\end{proof}

@@

\subsection{Identifiability}

In ordinary linear regression, the regression coefficients are identifiable
if and only if the data has full rank: $\E[x\tp{2}] \succ 0$.
However, in mixture of linear regressions, this is insufficient.  
For example, suppose $x = (1, t, \cdots, t^p)$, 
the common polynomial basis expansion.
Regressing on all powers up to $p$ yield at most $3$.

%$1, x_1, x_1^2, \dots, x_1^p$ is
%neither quadratically nor cubically independent because the product
%space is strictly a subset of $\Re^{\frac{d (d-1)}{2}}$ since it
%contains $x_1 \times x_1 = x_1^2 \times 1$, and $x_1 \times x_1 \times
%x_1 = x_1^2 \times x_1 \times 1 = x_1^3 \times 1 \times 1$.
%In fact, the
%number of unique terms in $x\tp{p}$ is only of the order of $p \times
%d$, and hence we will not have enough equations to solve for $M_2$ or
%$M_3$ unless we consider $p = d^2$ moments.

%The mixture of linear regressions model is identifiable subject to
%permutations of the components. However, our approach
%employs regression on $x\tp{2}$ and $x\tp{3}$, and thus requires
%$\mathcal{X} \otimes \mathcal{X}$ and $\mathcal{X} \otimes \mathcal{X}
%\otimes \mathcal{X}$ to be linearly independent modulo symmetry. We call
%these condition {\em quadratic} and {\em cubic} independence.

%This condition trivially holds when the space $\mathcal{X}$ consists of
%independent vectors $\{ x_1, x_2, \cdots, x_d \}$. Unfortunately, it
%places some constraints on non-linear featurizations.

### 2013-02-15 doc/icml2013/regression.tex
checked in
many changes

\section{Recovering $M_2, M_3$ with low-rank regression}
\label{sec:regression}

In this section, we analyze step 1 of the Spectral Experts
algorithm: that of recovering the compound parameters $M_2,M_3$
from data.

@@

\cite{candesPhaseLift} regresses on the square of the response,
because that's the observation.

Analysis for regression in the fixed and random design settings have
been studied before \citep{HsuKakadeZhang}, however our setup differs
substantially from the noise models assumed in the literature. In our
scenario, the variance in the estimation comes not only from the
Gaussian observation noise (which has been studied before), but also
from the variance in the latent variable $h$.

Let us now formally define the class of regression problems we wish to
analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
\begin{align*}
  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
\end{align*}
Note that the term in brackets is the error in our estimates and has
zero mean.  A more succinct way of representing this problem is to use
the operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{N}$, which uses the
input $\{ x_i \}_{i=1}^{N}$ to map $M_p$ to a vector of responses $\hat
y_i = \opX(M_p)_i = \innerp{x_i\tp{p}}{M_p}$. We can also represent the
errors with the vector $\vec \epsilon$. Rewriting the equation with this notation, 
\begin{align*}
  \vec y^p &= \opX(M_p) + \vec \epsilon.
\end{align*}

\todo{Describe/define the convex tensor stuff.} We would like to exploit
the property that $M_p$ is low rank (as typically $K \ll D$). It has
been shown that a convex relaxation for this problem regression with
trace norm regularization, which can be solved using a proximal
subgradient descent algorithm\citationneeded. The analogue of trace norm
regularization for higher order tensors corresponds to the sum of the
trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
by concatenating the entries for all dimensions other than $k$. For
example, the 1-mode unfolding of a 3rd order tensor has entries,
$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.

In general, the optimization problem we'd like to solve is,
\begin{align}
  \hat M_p &= 
  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
\end{align}

We begin our analysis by stating a generic theorem about convex
optimization problem in \equationref{eq:regression}, 
\begin{lemma}(Convergence of Trace-Norm Regularized Regression)
  Let $\hat M_p$ be the solution of the optimization problem
  \eqref{eq:regression} with $\lambda_N \ge 2 \frac{\| \opX^*(\epsilon)
  \|}{N}$. Suppose $\opX$ satisfies the restricted strong convexity
  condition, $$\frac{1}{2N} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX)
  \|\Delta\|^2_F$$, then
  $\| \hat M_p - M_p^* \|_F \le \frac{32 \lambda_N}{\kappa(\opX)} \frac{1}{K} \sum_{k=1}^{K} \sqrt{r_k}$.
\end{lemma}
\begin{proof}
  \citet{Tomioka2011}, Theorem 1.

@@

>>>>

\subsection{Low-rank Regression}
\label{sec:regression}

The goal of this section is to bound the error of
the compound parameter estimates,
specifically $\|\Delta_2\|_F^2$ and $\|\Delta_3\|_F^2$,
where $\Delta_2 \eqdef \hat M_2 - M_2$
and $\Delta_3 \eqdef \hat M_3 - M_3$.

Our analysis is based on the low-rank regression framework of
\citet{Tomioka2011} for tensors, which builds off of
\citet{NegahbanWainwright2009} for matrices.
Define the observation operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
mapping compound parameters $M_p$:
\begin{align}
  \opX(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
\end{align}

Let $\kappa(\opX)$ be the restricted strong constant,
and let $\opX^*(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
be the adjoint.

\paragraph{Restricted strong convexity}

In this section, we compute the restricted strong convexity constant
$\kappa(\opX)$ for our problem.

@@

%Analysis for regression in the fixed and random design settings have
%been studied before \citep{HsuKakadeZhang}, however our setup differs
%substantially from the noise models assumed in the literature. In our
%scenario, the variance in the estimation comes not only from the
%Gaussian observation noise (which has been studied before), but also
%from the variance in the latent variable $h$.

%Let us now formally define the class of regression problems we wish to
%analyze, i.e. regression on the set $(x\tp{p}, y^p)$,
%\begin{align*}
%  y^p &= \innerp{x\tp{p}}{M_p} + (\innerp{x\tp{p}}{M_p - \beta_h\tp{p}} + \varepsilon).
%\end{align*}

%\todo{Describe/define the convex tensor stuff.}
%We would like to exploit
%the property that $M_p$ is low rank (as typically $K \ll D$). It has
%been shown that a convex relaxation for this problem regression with
%trace norm regularization, which can be solved using a proximal
%subgradient descent algorithm\citationneeded.
%The analogue of trace norm
%regularization for higher order tensors corresponds to the sum of the
%trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
%$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
%by concatenating the entries for all dimensions other than $k$. For
%example, the 1-mode unfolding of a 3rd order tensor has entries,
%$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.

%In general, the optimization problem we'd like to solve is,
%\begin{align}
%  \hat M_p &= 
%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
%\end{align}

%We begin our analysis by stating a generic theorem about convex
%optimization problem in \equationref{eq:regression}, 
\begin{lemma}[convergence of low-rank regression]
If restricted strong convexity holds
$$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{|\opX^*(\eta_p)|}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX)}$.
\end{lemma}
\begin{proof}
  \citet{Tomioka2011}, Theorem 1.

@@

By Lemma 19 of \cite{hsu13spherical},
we have that $\frac{1}{n} \sum_{i=1}^n \epsilon_i^3 = O(\sqrt{\frac{\log^3(1/\delta)}{n}})$.
with probability at least $1-\delta$.
